---
title: "MSCA 31006 Time Series - Final Project - Holt Winters"
author: "Anna Auersperg"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# load packages
packages <- c("tidyverse","reshape2","xts","forecast","lubridate","tseries","fpp","tsibble","feasts","TSA","naniar","lubridate")
suppressPackageStartupMessages(lapply(packages, require, character.only = TRUE))

```

# Import and process data 

```{r}
# import data
raw.df <- read.csv("city_temperature.csv")

# remove observations with year = 200, 201 or 2020 (only partial data), Day = 0 
wip.df <- raw.df %>% 
  filter(!Year %in% (200:201)) %>%
  filter(Day != 0)

# date data type
wip.df <- wip.df %>%
  mutate(Date = as.Date(with(wip.df,paste(Year,Month,Day,sep="-")),
                        "%Y-%m-%d")) %>%
  select(-c("Month","Day","Year"))

# number of days between the start and end date (inclusive)
n_days <- max(wip.df$Date) - min(wip.df$Date) + 1
```

Subset US cities

```{r}
# list of US cities of interest
CitiesOfInterest <- c("Chicago",
                      "Anchorage",
                      "Denver",
                      "Indianapolis",
                      "Kansas City",
                      "Las Vegas",
                      "Los Angeles",
                      "Miami Beach",
                      "Minneapolis St. Paul",
                      "New York City",
                      "Phoenix",
                      "Raleigh Durham",
                      "San Antonio")

# subset US cities of interest 
df.usa <- wip.df %>% dplyr::filter((Country=="US") 
                                     & (City %in% CitiesOfInterest) 
                                     & (State!="Additional Territories")) %>%
  select(-c("Region", "Country", "State")) # drop redundant columns

# reshape data set, one column per city
df_melt <- melt(df.usa, id.vars = c("Date","City"))
df.usa2 <- dcast(df_melt, Date ~ City)

# update values where `AvgTemperature` = -99 by forward filling 
df.usa2 <- df.usa2 %>% naniar::replace_with_na_all(condition = ~.x == -99)
df.usa2 <- na.locf(na.locf(df.usa2), fromLast = FALSE)
```

## Univariate Time Series: Daily, Weekly, Monthly Avg Temp in Chicago

*Create Test and Train Splits*

- Train : 1995-01-01 - 2019-12-31
- Test  : 2020-01-01 - 2020-05-13

```{r}
train.test <- function(ts, year, split){
  if(split == 'train'){
    ts <- window(ts, end=c(year-1,12))
  }
  else{
    ts <- window(ts, start=c(year,1))
  }
  return(ts)
}
```

```{r}
# select Chicago
df <- df.usa2 %>% select(c("Date","Chicago")) %>% rename(AvgTemp = "Chicago")
```

### Function to generate time series of frequency specified 


```{r}
generateTS <- function(freq){
  if(freq == 'week'){
     # group by week, use week start date 
      ts <- df %>% group_by(WeekStart = lubridate::floor_date(df$Date,unit="week")) %>%
        summarize(AvgTemp = mean(AvgTemp) %>% round(2)) %>% 
        select(-c("WeekStart")) %>% ts(start=c(1995,1), frequency=52)
  }
  if(freq == 'month'){
    ts <- df %>% group_by(YearMonth = as.yearmon(Date)) %>%
      summarize(AvgTemp = mean(AvgTemp) %>% round(2)) %>%   
      select(-c("YearMonth")) %>%
      ts(start=c(1995,1), frequency=12)
  }
  else{
    ts <-  df %>% select(-c("Date")) %>% ts(start=c(1995,1), frequency=365)
  }
  ts 
}
```

# Model

```{r}

#accepts model predictions and test dataframe 
#idea here is we can merge the results for each frequency into one frame and change predicted vals to the model results per model 
EvaluateModel <- function(pred, test, freq){
  data.frame(Date = seq(from = as.Date("2020-01-01"), by = freq, length.out = length(test)),
                    Observed = as.numeric(test),
                    Predicted=pred$mean)
}
```
## Holt Winters
```{r}
HW.Mod <- function(freq){
  train <- generateTS(freq) %>% train.test(2020,"train")
  test <- generateTS(freq) %>% train.test(2020,"test")
  
  hw <- HoltWinters(train)
  print(hw)
  pred <- forecast(hw, h=length(test))
  print(plot(pred))
  EvaluateModel(pred, test, freq)
}
HW.fc <- lapply(c("day","week","month"),HW.Mod)
```

df %>%
  # group by YearMonth
  group_by(YearMonth = as.yearmon(Date)) %>%
  # calculate avg
  summarize(AvgTemp = mean(AvgTemp) %>% round(2)) %>% 
  # drop "YearMonth" date before converting to ts
  select(-c("YearMonth")) %>%
  # convert to time series object
  ts(start=c(1995,1), frequency=12)

